---
title: "Prediction_Assignment_Somik"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


##Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

##Loading Required Libraries

```{r}
library(knitr)
library(caret)
library(corrplot)
library(rattle)
library(doParallel)
registerDoParallel(detectCores()-1)
```

## Loading Required Data

```{r}
pml_data<-read.csv("~/R Coursera/pml-training.csv")
pml_predict_data<-read.csv("~/R Coursera/pml-testing.csv")
pml_data<-pml_data[,-1]
pml_predict_data<-pml_predict_data[,-1]
kable(head(pml_data))
kable(head(pml_predict_data))
```

Now we will spit the data into a Training and Test Data set

```{r}
inTrain<-createDataPartition(pml_data$classe,p=0.7,list = FALSE)
pml_training<-pml_data[inTrain,]
pml_testing<-pml_data[-inTrain,]
```

###Data Cleaning

From the Data we can see that there is are a lot of NAs i.e missing data in the dataset. We would need to clean it as it will skew the predictions made by our model if we include those predictor variables.

Before we remove the missing data, we also need to remove the Near Zero Variance Variables as there are many such variables with only one value or more than 90% of one value.

```{r}
NZV<-nearZeroVar(pml_training)
pml_training<-pml_training[,-NZV]
dim(pml_training)
```

For NAs, I have removed variables for which number of values having NAs is greater than 90% of the total row count.

```{r}
NA_Columns <-sapply(pml_training,function(x) sum(is.na(x)))>0.90*nrow(pml_training)
pml_training<-pml_training[,NA_Columns==FALSE]
```

## Checking Correlations

In order to proceed further,it is important to check whether the predictor variables are highly correlated with each other. In order to do that, I have used cor to find the correlation coefficient and corrplot to visualize the correlations. cor only takes a numerical matrix as input and hence, we need to remove teh categorical variables from our data. 

```{r fig.height=12, fig.width=12}
str(pml_training)
corMatrix <- cor(pml_training[, c(-58,-1,-4)])
corrplot(corMatrix,method = "square",type = "upper")
```

As we can see, the variables are very low to moderately correlated and hence we can go ahead with all of the predictor variables to build our model.


## Modelling

For this analysis, I would be using three ML algorithms, Basic Tree model, Random Forest and Gradient Boosting Model. 

###CART MODEL

```{r}
tree_model<-train(classe~.,data=pml_training,method="rpart")
fancyRpartPlot(tree_model$finalModel)
```

###Random Forest

We will use K-fold cross validation method to build this model with 3 samples.

```{r}
rfcontrol<-trainControl(method="cv",allowParallel = TRUE,number = 3)
RF_model<-train(classe~.,data=pml_training,method="rf",trControl=rfcontrol)
print(RF_model)
plot(RF_model$finalModel)
```

###Gradient Boosting Method

We will use K-fold cross validation method to build this model with 5 samples.

```{r}
gbControl<-trainControl(method = "cv",allowParallel = TRUE, number = 5)
GBM_model<-train(classe~.,data=pml_training,method="gbm",trControl=gbControl,verbose=FALSE)
print(GBM_model)
```


##Prediction using Test Data and confusion Matrix

```{r}
predict_tree<-predict(tree_model,pml_testing)
predict_rf<-predict(RF_model,pml_testing)
predict_GBM<-predict(GBM_model,pml_testing)
```

```{r}
confusionMatrix(predict_tree,pml_testing$classe)
confusionMatrix(predict_rf,pml_testing$classe)
confusionMatrix(predict_GBM,pml_testing$classe)
```

Based on the confusion matrix, we can see that the Random forest has the highest accuracy with 99.95% closely followed by GBM model which has an accuracy of 99.69%. The basic tree model has the lowest accuracy of 49%.

Hence, Random Forest model would be the obvious choice amongst the three models.

## Prediction

```{r}
prediction<-predict(RF_model,pml_predict_data)
prediction
```

